### 2018-05-29 4_impl.tex
add pictures and stuff.

Secondly, \porthos{} resolves the semantics of operations syntactically (it was defined in the \tool{ANTLR} grammar), whereas it should be resolved by a separate module operating on the AST level%
%
%TODO: check or remove this
%\footnote{Most modern languages (including C++) allow the function overloading, thus requiring the type information to be available for the function resolution algorithm (see Section~\ref{ch:impl:proc:x-pre-compiler}).}%
%
, so that it does not require to change grammar for encoding the semantics of a new function. %adding support for a new function.
%\footnote{for instance, given two functions `\lstinline{foo(int a)}' and `\lstinline{foo(char a)}' compiled by a C++ compiler, the code `\lstinline{int a = '1'; foo(a);}' will invoke the first method rather than the second one.}.%
As the reader might have noticed from the grammar sketch in Figure~\ref{fig:in_grammar_pts}, different kinds of memory operations of of the \oldporthos{} input language vary syntactically as well.
For example, the assignment of a local computation to a register uses the symbol `\lstinline{<-}', the atomic non-relaxed load operation is denoted by \lstinline{<:-}', atomic non-relaxed store operation is denoted by `\lstinline{:=}', and the semantics of relaxed \lstinline{load} and \lstinline{store} is resolved syntactically by matching the function name and arguments.
Moreover, only the operator `\lstinline{<-}' accepts an expression as the source of data, which means that expressions could be assigned only to registers.
In \porthos[2], the semantics of the data-flow operation is determined according to the types of operands, that are determined during the pre-compilation stage (see Section~\ref{ch:impl:proc:x-pre-compiler}).
The semantics of the functions also being resolved during the compilation stage via the \textit{invocation hooking} mechanism (see Section~\ref{ch:impl:proc:x-compiler}).

@@

The general processing scheme of old tool \porthos[1] is the following.
The input program is parsed to the AST, where leaves are the events.
The instructions of the AST are then encoded into the $\phi_{CF}$ (see Section~\ref{ch:enc:bmc:cf}) and the events are processed by the single function (`\texttt{dartagnan.wmm.Domain.encode}') that loops over all events, checks their properties (the kind of event, the operands of the memory event, the type of the barrier event, etc.) and adds corresponding relations assertions to the formula.
Encoding all the relations (the \textit{program domain encoding}) in a single function is error-prone and hardly manageable.
Moreover, this function does not consider the memory model, therefore the result formula can contain relations that are not constrained by the memory model specified by the user and thus are redundant (for instance, the \texttt{TSO} memory model does not contain \texttt{POWER}-specific relations \texttt{ii}, \texttt{ic}, \texttt{ci}, etc., that are always encoded regardless the memory model).
In \porthos[2], we moved the encoding logic into a separate module \ref{ch:impl:proc:z-encoder} that accepts the input program and the user-defined memory model, therefore does not add constraints over unused relations to the result SMT-formula.

The high-level architecture of \porthos[2] is presented in Figure~\ref{fig:arch}.
The program takes as input the program to be analysed and one (the reachability analysis mode) or two (the portability analysis mode) memory models.
However, instead of parsing memory models from the \cat{} file, the tool may operate with pre-defined memory models (\texttt{SC}, \texttt{TSO}, \texttt{PSO}, \texttt{RMO}, \texttt{ALPHA}, \texttt{POWER}, \texttt{ARM}; this feature is inherited from \porthos[1]).
The parsed program syntax tree is then converted (Section~\ref{ch:impl:proc:y-constr}) to a program AST called \ytree{}%
%
\footnote{In order to avoid confusion between different internal representations, we prefix the names of elements of each internal representation with a letter. For instance, we picked the letter `Y' to denote the AST code representation as drawing of this letter resembles the tree branching; with letter `X' we prefix elements of the event-flow graph as the events are to be e\textbf{x}ecuted; and with letter `W' we prefix elements of the \textbf{w}eak memory model AST.}%
%
(Section~\ref{ch:impl:model:ytree}), which then is being preprocessed at the pre-compilation stage (Section~\ref{ch:impl:proc:x-pre-compiler}) in order to collect information necessary for the compilation.
The \ytree{} then is being compiled (Section~\ref{ch:impl:proc:x-compiler}) to an \xgraph{} representation (Section~\ref{ch:impl:model:xgraph}).
The compiled \xgraph{} then is being converted to an acyclic form (Section~\ref{ch:impl:proc:x-unroll}) in order to be encoded into a \zformula{} (Section~\ref{ch:impl:model:zformula}).
Apart from that, the memory-model constructor (Section~\ref{ch:impl:proc:w-constr}) constructs the abstract syntax tree of derived relations of the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
%on the basis of the \xgraph{} in order to build the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT-formula), which then is translated to an SMT-formula (Section~\ref{ch:impl:proc:smt-converter}), which then is solved by the SMT-solver (Section~\ref{ch:impl:proc:smt-solver}).

@@

The \ytree{} is compiled into the low-level event-based program representation called \textit{\xgraph{}}.
The mathematical structure of event-flow graph was discussed in Section~\ref{ch:wmm:model}.
The nodes of the graph are events, and the edges are basic relations: the control-flow relation \po{} and the data-flow relations \co{} and \rf{}.
Hereinafter, we denote the \xgraph{} with only control-flow edges as \xgraph[CF], the \xgraph{} with only data-flow edges as \xgraph[CF].
The complete \xgraph{} is $\xgraph[CF+DF] = \xgraph[CF] \; \cup \; \xgraph[DF]$.
The UML diagram on Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the hierarchy of main interfaces of the X-abstraction level.
The \textit{full} list of classes that constitute the \xgraph{} hierarchy can be found in Appendix~\ref{apx:trees}.
Internally, the graph is represented by an adjacency matrix (to be exact, by multiple adjacency matrices that store edges of different kinds, see more details in Section~\ref{ch:impl:model:xgraph}).

@@

Following the terminology of the C standard, we distinguish the \textit{r-value} and \textit{l-value} memory units (unlike r-values, the l-values may be assigned a new value).
As r-values cannot change their value, they can be seen as the value itself (therefore the \texttt{XComputationEvent} is modelled as an local r-value memory unit, see more detailed discussion further in current Section).

@@

The main class representing the X-compiler is \texttt{Y2XConverter}. %TODO: draw it on Figure.
It receives as input the \ytree{}, the memory model kind and user settings (for instance, the interpreter mode defining the set of invocation hooks enabled during the analysis run).
The \texttt{Y2XConverter} creates an instance of the stateless visitor \texttt{Y2XConverterVisitor} that traverses the \ytree{} while invoking the stateful interpreter \texttt{XInterpreter}.
The \texttt{XInterpreter} carries the \xgraph{}-builder and provides \textit{action} methods for changing its state and thus the filling in the builder.
The interpreter has additional modules \texttt{XMemoryManager}, \texttt{XHookManager} and \texttt{XTypeManager} that provide specialised functionality.

@@

The unrolling procedure is performed by the \texttt{XFlowGraphUnroller}.
For unrolling the flow-graph, we perform the \textit{Deep-First Search~(DFS)} while counting the X-events (for the unrolling bound) and keeping track of the depth stack (for detecting back edges needed for determining the type of sink node).
Back-edges are determined via the depth stack (each next event is pushed onto the stack; if the new event is present in the stack, the last processed edge is a back-edge).
Also, during the unrolling each next event increments the \textit{unrolling depth counter} (the \textit{event reference-ID}) -- the integer stored by the \texttt{XEvent} thus making it an \textit{event reference} on non-zero values (only events of a non-unrolled graph can have the zero-valued reference-ID).
As an X-event is immutable, it has the method `\texttt{XEvent asNodeRef(int refId)}' that clones it with new value of reference-ID.
The methods \texttt{hashCode} and \texttt{equals} consider the reference-ID as the uniqueness field for all events except sink and source events (considering the usage of \texttt{HashMap} and Guava's \texttt{ImmutableMap} for storing events, proper setting of the hash-code methods is a crucial programming task).

>>>>

Secondly, \porthos{} resolves the semantics of operations syntactically (it was defined in the \tool{ANTLR} grammar), whereas it should be resolved by a separate module operating on the AST level, %
%
%TODO: check or remove this
%\footnote{Most modern languages (including C++) allow the function overloading, thus requiring the type information to be available for the function resolution algorithm (see Section~\ref{ch:impl:proc:x-pre-compiler}).}%
%
so that it does not require to change the grammar for encoding the semantics of a new function. %adding support for a new function.
%\footnote{for instance, given two functions `\lstinline{foo(int a)}' and `\lstinline{foo(char a)}' compiled by a C++ compiler, the code `\lstinline{int a = '1'; foo(a);}' will invoke the first method rather than the second one.}.%
As the reader might have noticed from the grammar sketch in Figure~\ref{fig:in_grammar_pts}, different kinds of memory operations of of the \oldporthos{} input language vary syntactically as well.
For example, the assignment of a local computation to a register uses the symbol `\lstinline{<-}', the atomic non-relaxed load operation is denoted by \lstinline{<:-}', atomic non-relaxed store operation is denoted by `\lstinline{:=}', and the semantics of relaxed \lstinline{load} and \lstinline{store} is resolved syntactically by matching the function name and arguments.
Moreover, only the operator `\lstinline{<-}' accepts an expression as the source of data, which means that expressions could be assigned only to registers.
In \porthos[2], the semantics of the data-flow operation is determined according to the types of operands, that are determined during the pre-compilation stage (see Section~\ref{ch:impl:proc:x-pre-compiler}).
The semantics of the functions also being resolved during the compilation stage via the \textit{invocation hooking} mechanism (see Section~\ref{ch:impl:proc:x-compiler}).

@@

The general processing scheme of old tool \porthos[1] is the following.
The input program is parsed to the AST, where leaves are the events.
The instructions of the AST are then encoded into the $\phi_{CF}$ (see Section~\ref{ch:enc:bmc:cf}) and the events are processed by the single function (`\texttt{dartagnan.wmm.Domain.encode}') that loops over all events, checks their properties (the kind of event, the operands of the memory event, the type of the barrier event, etc.) and adds corresponding relations assertions to the formula.
Encoding all the relations (the \textit{program domain encoding}) in a single function is error-prone and hardly manageable.
Moreover, this function does not consider the memory model, therefore the result formula can contain relations that are not constrained by the memory model specified by the user and thus are redundant (for instance, the \texttt{TSO} memory model does not contain \texttt{POWER}-specific relations \texttt{ii}, \texttt{ic}, \texttt{ci}, etc., that are always encoded regardless the memory model).
In \porthos[2], we moved the encoding logic into a separate module (see Section~\ref{ch:impl:proc:z-encoder}) that accepts the input program and the user-defined memory model, therefore does not add constraints over unused relations to the result SMT-formula.

The high-level architecture of \porthos[2] is presented in Figure~\ref{fig:arch}.
The program takes as input the program to be analysed and one (the reachability analysis mode) or two (the portability analysis mode) memory models.
However, instead of parsing memory models from the \cat{} file, the tool may operate with pre-defined memory models (\texttt{SC}, \texttt{TSO}, \texttt{PSO}, \texttt{RMO}, \texttt{ALPHA}, \texttt{POWER}, \texttt{ARM}; this feature is inherited from \porthos[1]).
The parsed program syntax tree is then converted (Section~\ref{ch:impl:proc:y-constr}) to a program AST called \ytree{}%
%
\footnote{In order to avoid confusion between different internal representations, we prefix the names of elements of each internal representation with a letter. For instance, we picked the letter `Y' to denote the AST code representation as drawing of this letter resembles the tree branching; with letter `X' we prefix elements of the event-flow graph as the events are to be e\textbf{x}ecuted; and with letter `W' we prefix elements of the \textbf{w}eak memory model AST.}%
%
(Section~\ref{ch:impl:model:ytree}), which then is being preprocessed at the pre-compilation stage (Section~\ref{ch:impl:proc:x-pre-compiler}) in order to collect information necessary for the compilation.
The \ytree{} then is being compiled (Section~\ref{ch:impl:proc:x-compiler}) to an \xgraph{} representation (Section~\ref{ch:impl:model:xgraph}).
The compiled \xgraph{} then is being converted to an acyclic form (Section~\ref{ch:impl:proc:x-unroll}) in order to be encoded into a \zformula{} (Section~\ref{ch:impl:model:zformula}).
Apart from that, the memory-model constructor (Section~\ref{ch:impl:proc:w-constr}) constructs the abstract syntax tree of derived relations of the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
%on the basis of the \xgraph{} in order to build the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT-formula), which then is translated to an SMT-formula, which then is solved by the SMT-solver.
Although this general architecture is defined for

@@

The \ytree{} is compiled into the low-level event-based program representation called \textit{\xgraph{}}.
The mathematical structure of event-flow graph was discussed in Section~\ref{ch:wmm:model}.
The nodes of the graph are events, and the edges are basic relations: the control-flow relation \po{} and the data-flow relations \co{} and \rf{}.
Hereinafter, we denote the \xgraph{} with only control-flow edges as \xgraph[CF], the \xgraph{} with only data-flow edges as \xgraph[CF].
The complete \xgraph{} is $\xgraph[CF+DF] = \xgraph[CF] \; \cup \; \xgraph[DF]$.
The UML diagram in Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the class hierarchy%
%
\footnote{Hereinafter all class diagrams are generated by IntelliJ IDEA Ultimate.}%
%
of main interfaces of the X-abstraction level.
The \textit{full} list of classes that constitute the \xgraph{} hierarchy can be found in Appendix~\ref{apx:trees}.
Internally, the graph is represented by an adjacency matrix (to be exact, by multiple adjacency matrices that store edges of different kinds, see more details in Section~\ref{ch:impl:model:xgraph}).

@@

Following the terminology of the C standard, we distinguish the \textit{r-value} and \textit{l-value} memory units (unlike r-values, the l-values may be assigned a new value).
As r-values cannot change their value, they can be seen as the value itself (therefore the \texttt{XComputationEvent} is modelled as an local r-value memory unit, see more detailed discussion further in the current Section).

@@

The main class representing the X-compiler is \texttt{Y2XConverter}. %TODO: draw it in Figure.
It receives as input the \ytree{}, the memory model kind and user settings (for instance, the interpreter mode defining the set of invocation hooks enabled during the analysis run).
The \texttt{Y2XConverter} creates an instance of the stateless visitor \texttt{Y2XConverterVisitor} that traverses the \ytree{} while invoking the stateful interpreter \texttt{XInterpreter}.
The \texttt{XInterpreter} carries the \xgraph{}-builder and provides \textit{action} methods for changing its state and thus the filling in the builder.
The interpreter has additional modules \texttt{XMemoryManager}, \texttt{XHookManager} and \texttt{XTypeManager} that provide specialised functionality.

@@

The unrolling procedure is performed by the \texttt{XFlowGraphUnroller}.
For unrolling the flow-graph, we perform the \textit{Deep-First Search~(DFS)} while counting the X-events (for the unrolling bound) and keeping track of the depth stack (for detecting back edges needed for determining the type of sink node).
Back-edges are determined via the depth stack (each next event is pushed onto the stack; if the new event is present in the stack, the last processed edge is a back-edge).
Also, during the unrolling each next event increments the \textit{unrolling depth counter} -- the non-zero integer \textit{event reference-ID} stored by each X-event (only events of a non-unrolled graph can have the zero-valued reference-ID)
As an X-event is immutable, it has the method `\texttt{XEvent asNodeRef(int refId)}' that clones it with new value of reference-ID.
The methods \texttt{hashCode} and \texttt{equals} consider the reference-ID as the uniqueness field for all events except sink and source events (considering the usage of \texttt{HashMap} and Guava's \texttt{ImmutableMap} for storing events, proper setting of the hash-code methods is a crucial programming task).

### 2018-05-29 5_eval.tex
add pictures and stuff.

\subsection{Compilation and unrolling}
\label{ch:eval:show:compil}

%As an example of the compilation and unrolling processes, consider the simple
%As a simple example of the compilation process discussed in Section~\ref{ch:impl:proc:x-compiler}, 
Consider two equivalent functions in C (left) and the \porthos[1] input language (right) represented in Figure~\ref{ex:test-both-cf}. %(in the C language on the left-hand side, the \porthos[1] input language on the right-hand side).
%The functions are state-equivalent (represent the same state machine) and 
The functions does not compute any useful value, however they contain three nested \texttt{while}-loops and thus can serve as an illustration of the differences in the program compilation and unrolling between \porthos[1] and \porthos[2].
%contains several syntactic elements (such as prefix increment and loop-breaking statements), which are supported by \porthos[2] comparing to its predecessor.
%In the Figure, the control-flow subgraph \xgraph[CF] of the non-unrolled event-flow graph is presented on the right-hand side (

\begin{figure}[!h]
%\begin{minipage}{.53\textwidth}
\begin{subfigure}[b]{.53\textwidth}\centering
\begin{lstlisting}[language=Java,basicstyle=\ttfamily\small]

@@

\end{lstlisting}
\caption{An example in the C language}
\label{ex:both-cf:ptsC}
\end{subfigure}
%\begin{minipage}{.45\textwidth}
\begin{subfigure}[b]{.45\textwidth}\centering
\begin{lstlisting}[language=Java,basicstyle=\ttfamily\small]
{ x }
thread t0 {
     a <- 1;
     c <- 1;
     while (a == 1) {
         b <- 1;
         while (b == 1) {
             while (c == 1) {
                 x := c
             };
             x := b
         };
     };
     x := a
 }
\end{lstlisting}
\caption{An example in the \porthos[1] input language}
\label{ex:both-cf:pts1}
\end{subfigure}
\caption{Example: A demonstrative cyclic function}
\label{ex:test-both-cf}
\end{figure}

The functions are processed by \porthos[2] or modified version \porthos[1] that is able to print same \xgraph[CF] as the new tool (for that, the control-flow instructions \texttt{if-then-else}, \texttt{while} and \texttt{sequence} are expanded recursively and the head and the tails of each instruction are bind by the method processing its parent).
The graph generation is performed via the open-source library \texttt{Graphviz}~\cite{ellson2001graphviz}).
%Each event of the control-flow graph contains the unique number (generated by the \texttt{hashCode} method) in curly brackets below the value, which is necessary for correct displaying the graph.

The Figure~\ref{ex:test-both-pic} illustrates the data structure which the functions in Figure~\ref{ex:test-both-cf} are compiled to.
The left-hand side picture represents the non-unrolled \xgraph[CF] generated by \porthos[2], and the right-hand picture represents the AST generated by \porthos[1].

In both pictures, the writes are denoted with the left-directed arrow `\lstinline{<-}', and the functions \lstinline{load} and \lstinline{store} denote the type of the shared memory event.
The primary transitions that denote unconditional jumps or if-true-transitions are pictured with solid lines, and the alternative transitions that denote if-false-transitions are pictured with dotted lines.
The graphs contains a single source event and a single sink event represented by the grey triangles (actually, the graph produced by \porthos[1] does not have sink and source nodes, but they were added to the picture for demonstrative purposes).
For clarity, all branching events, that in current example serve as the conditional events of loops, are highlighed with light-grey colour.

\begin{figure}[!h]
%
\begin{subfigure}[t]{.49\textwidth}\centering
  \includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/unrolling-comparison/PorthosC/t0.png}
  \hfill
  \caption{The compiled \xgraph{} of the function in Figure~\ref{ex:both-cf:ptsC}}
  \label{ex:both-cf:graph:ptsC}
\end{subfigure}
%
\begin{subfigure}[t]{.49\textwidth}\centering
  \includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/unrolling-comparison/Porthos/t0.png}
  \hfill
  \caption{The compiled AST of the function in Figure~\ref{ex:both-cf:pts1}}
  \label{ex:both-cf:graph:pts1}
\end{subfigure}
%
\caption{The control-flow graphs of the functions represented in Figure~\ref{ex:test-both-cf}}
\label{ex:test-both-pic}
\end{figure}

Two compiled the graphs are equivalent up to the extra \texttt{nop}-events in the \porthos[2] graph, that are necessary for correct encoding as it was discussed in Section~\ref{ch:enc:bmc:cf}, and \texttt{skip}-events in \porthos[1] graph.
However, the unrolled graphs presented in Figure~\ref{ex:test-both-pic-unroll} are different as \porthos[2] uses different unrolling algorithm.
The labels of events in the left-hand side picture (produced by \porthos[2]) are augmented by the unrolling depth number, which is separated from the event label by comma.

The unrolling algorithm used by \porthos[1] (right-hand side picture) unrolls \textit{all} loops $k$ times (where $k$ is the unrolling bound), and the unrolling algorithm of \porthos[2] unrolls loops so that not more than $k$ events are executed.
As it is illustrated by the picture, the new algorithm produces a better set of program executions (for example, the unrolled graph of \porthos[1] does not contain executions of the inner loops more than $k=2$ times, which makes the old unrolling algorithm not complete).
As the new unrolling algorithm uses the Deep-First Search, it discovers \textit{all} possible paths, therefore the result graph contains \textit{all} possible executions and thus is complete.

@@

\begin{figure}[!h]
%
\begin{subfigure}[b]{.6\textwidth}\centering
  \includegraphics[height=.6\textheight,width=\textwidth]{img/my/graphs/unrolling-comparison/PorthosC/t0_unrolled.png}
  \hfill
  \caption{The unrolled \xgraph{} of the function in Figure~\ref{ex:both-cf:ptsC}}
  \label{ex:both-cf:graphU:ptsC}
\end{subfigure}
\hfill
%
\begin{subfigure}[b]{.3\textwidth}\centering
  \includegraphics[height=.9\textheight,width=\textwidth]{img/my/graphs/unrolling-comparison/Porthos/t0_unrolled.png}
    \caption{The unrolled AST of the function in Figure~\ref{ex:both-cf:pts1}}
    \hfill
  \label{ex:both-cf:graphU:pts1}
\end{subfigure}
\hfill
%
\caption{The unrolled control-flow graphs of the functions represented in Figure~\ref{ex:test-both-cf}}
\label{ex:test-both-pic-unroll}
\end{figure}

@@

%TODO: here (or better in Chap 5 ) say about 'we tried different encoding schemes !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

@@

%TO-DO: fix interface
%To run \porthos[2] in reachability analysis mode, we use the command `\lstinline{mousq

%\textbf{TODOOOOOOOOOOOOOOOOOOOOO!!!! Re-compute tests with new counting of encoding}
When running in the state reachability analysis mode, \porthos[2] produces the following output:

\begin{lstlisting}
$ java PorthosC --reachability --input benchmarks/C11/Dekker.c \
      --bound 27 --target TSO
0.288: Interpretation...
0.426: Unrolling...
0.533: Program encoding...
0.701: Program domain encoding...
#=193 (107)
34.187: Memory model encoding...
36.818: Solving...
{
  "result": "NonReachable",
  "mainTimer": {
    "elapsedTimeSec": 36.163
  },
  "timers": {
    "Interpretation": {
      "elapsedTimeSec": 0.132
    },
    "ProgramDomainEncoding": {
      "elapsedTimeSec": 32.486
    },
    "Solving": {
      "elapsedTimeSec": 0.345
    },
    "Unrolling": {
      "elapsedTimeSec": 0.061
    },
    "MemoryModelEncoding": {
      "elapsedTimeSec": 2.631
    },
    "ProgramEncoding": {
      "elapsedTimeSec": 0.168
    }
  },
  "errors": []
}
\end{lstlisting}

@@

@@

<TODO: above as table>

@@

<TODO: profile both programs, say sth about memory usage!>

@@

For evaluating the tool working in the portability analysis mode, we used the same file \texttt{Dekker.c} tested for the state-portability from TSO to SC: %TODO: Say sth about state-portability

%!!! \textit{\textbf{TODO: change the main class name to PorthosC }} !!!

\begin{lstlisting}
$ java PorthosC --portability --input benchmarks/C11/Dekker.c \
      --bound 27 --source TSO --target SC --state
0.359: Interpretation...
0.364: Unrolling...
0.590: Unrolling...
0.623: Program encoding...
0.847: Program domain encoding...
#=193 (107)
#=193 (107)
66.541: Memory model encoding...
72.321: Solving...
{
  "result": "StatePortable",
  "iterations": 0,
  "mainTimer": {
    "elapsedTimeSec": 73.02
  },
  "timers": {
    "ProgramEncoding": {
      "elapsedTimeSec": 0.223
    },
    "MemoryModelEncoding": {
      "elapsedTimeSec": 5.779
    },
    "Interpretation": {
      "elapsedTimeSec": 0.248
    },
    "Solving": {
      "elapsedTimeSec": 0.699
    },
    "ProgramDomainEncoding": {
      "elapsedTimeSec": 65.693
    },
    "Unrolling": {
      "elapsedTimeSec": 0.022
    }
  },
  "errors": []
}
\end{lstlisting}

As the portability analysis requires compiling the program under two memory models, the overall program domain encoding time 65.693 sec is almost double as the same time in reachability analysis mode (\porthos[1] shows program domain encoding time 41.027 sec with the bound $k_1=3$).

>>>>

Consider two equivalent functions in C (left) and the \porthos[1] input language (right) represented in Figure~\ref{ex:test-both-cf}.
The functions contain three nested \texttt{while}-loops and thus can serve as an illustration of differences in the program compilation and unrolling processes between \porthos[1] and \porthos[2].

\begin{figure}[!h]
\begin{subfigure}[b]{.53\textwidth}\centering
\begin{lstlisting}[language=Java,basicstyle=\ttfamily\small]

@@

\end{lstlisting}
\caption{An example in the C language}
\label{ex:both-cf:ptsC}
\end{subfigure}
\begin{subfigure}[b]{.45\textwidth}\centering
\begin{lstlisting}[language=Java,basicstyle=\ttfamily\small]
{ x }
thread t0 {
     a <- 1;
     c <- 1;
     while (a == 1) {
         b <- 1;
         while (b == 1) {
             while (c == 1) {
                 x := c
             };
             x := b
         };
     };
     x := a
 }
\end{lstlisting}
\caption{An example in the \porthos[1] input language}
\label{ex:both-cf:pts1}
\end{subfigure}
\caption{Example: A demonstrative cyclic function}
\label{ex:test-both-cf}
\end{figure}

The functions are analysed for reachability of the final state by \porthos[2] and modified version \porthos[1] which is able to render the program AST to an event-based control-flow graph (for that, the control-flow instructions \texttt{if-then-else}, \texttt{while} and \texttt{sequence} are expanded recursively and the head and the tails of each instruction are bound by the method processing the parent).
The graph generation is performed via the open-source library \texttt{Graphviz}~\cite{ellson2001graphviz}).
%Each event of the control-flow graph contains the unique number (generated by the \texttt{hashCode} method) in curly brackets below the value, which is necessary for correct displaying the graph.

\subsection{Compilation}
\label{ch:eval:show:compil}

Figure~\ref{ex:test-both-pic} illustrates the data structure which the functions in Figure~\ref{ex:test-both-cf} are compiled to.
The left-hand side picture represents the non-unrolled \xgraph[CF] generated by \porthos[2], and the right-hand picture represents the AST generated by \porthos[1].

In both pictures, the writes are denoted with the left-directed arrow `\lstinline{<-}', and the functions \lstinline{load} and \lstinline{store} denote the type of the shared memory event.
The primary transitions that denote unconditional jumps or if-true-transitions are pictured with solid lines, and the alternative transitions that denote if-false-transitions are pictured with dotted lines.
The graphs contain a single source event and a single sink event represented by the grey triangles (in fact, the graph produced by \porthos[1] does not have sink and source nodes, but they were added to the picture for demonstrative purposes).
For clarity, all branching events, that in current example serve as the conditional events of loops, are highlighed with light-grey colour.

\begin{figure}[!h]
%
\begin{subfigure}[b]{.49\textwidth}\centering
  \includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/unrolling-comparison/PorthosC/t0.png}
  \hfill
  \caption{The compiled \xgraph{} of the function in Figure~\ref{ex:both-cf:ptsC}}
  \label{ex:both-cf:graph:ptsC}
\end{subfigure}
%
\begin{subfigure}[b]{.49\textwidth}\centering
  \includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/unrolling-comparison/Porthos/t0.png}
  \hfill
  \caption{The compiled AST of the function in Figure~\ref{ex:both-cf:pts1}}
  \label{ex:both-cf:graph:pts1}
\end{subfigure}
%
\caption{The control-flow graphs of the functions represented in Figure~\ref{ex:test-both-cf}}
\label{ex:test-both-pic}
\end{figure}

Note that two compiled graphs are equivalent up to the extra \texttt{nop}-events in the \porthos[2] graph, that are necessary for correct encoding as it was discussed in Section~\ref{ch:enc:bmc:cf}, and \texttt{skip}-events in \porthos[1] graph.
However, the unrolled graphs presented in Figure~\ref{ex:test-both-pic-unroll} are different as the tools use different unrolling algorithms.
The labels of events in the left-hand side picture (produced by \porthos[2]) are augmented by the unrolling depth number, which is separated from the event label by comma.

\subsection{Unrolling}
\label{ch:eval:show:unrol}

The unrolling algorithm used by \porthos[1] (right-hand side picture) unrolls \textit{all} loops $k$ times (where $k$ is the unrolling bound), and the unrolling algorithm of \porthos[2] unrolls loops so that not more than $k$ events are executed.
As it is illustrated by the picture, the new algorithm produces a better set of program executions (for example, the unrolled graph of \porthos[1] does not contain executions of the inner loops more than $k=2$ times, which makes the old unrolling algorithm not complete).
As the new unrolling algorithm is based on the DFS algorithm, it discovers \textit{all} possible paths, therefore the result graph contains \textit{all} possible executions and thus the program analysis performed on this graph can be complete up to bound~$k$.

@@

\begin{figure}%[!h]
%
\begin{subfigure}[b]{.6\textwidth}\centering
  \includegraphics[height=.65\textheight,width=1.1\textwidth]{img/my/graphs/unrolling-comparison/PorthosC/t0_unrolled.png}
  \hfill
  \caption{The unrolled \xgraph{} of the function in Figure~\ref{ex:both-cf:ptsC}}
  \label{ex:both-cf:graphU:ptsC}
\end{subfigure}
\hfill
%
\begin{subfigure}[b]{.3\textwidth}\centering
  \includegraphics[height=.9\textheight,width=\textwidth]{img/my/graphs/unrolling-comparison/Porthos/t0_unrolled.png}
    \caption{The unrolled AST of the function in Figure~\ref{ex:both-cf:pts1}}
    \hfill
  \label{ex:both-cf:graphU:pts1}
\end{subfigure}
\hfill
%
\caption{The unrolled control-flow graphs of the functions represented in Figure~\ref{ex:test-both-cf}}
\label{ex:test-both-pic-unroll}
\end{figure}

Figure~\ref{dep:events} illustrates the dependency of overall number of events in all unrolled event-flow graphs on the unrolling bound $k$ for the \texttt{Dekker} test.
%The left-hand side picture presents the dependency graph for \porthos[2], 
Both charts are limited above by $420$ events.
\porthos[2] (left-hand chart) reaches this limit with the bound $k_2=32$, and \porthos[1] (right-hand chart) reaches it with $k_1=5$.
However, taking into account the fact that initially both non-unrolled graphs had $20$ events, \porthos[2] at the step $32$ (the longest execution) obtained same number of events  as \porthos[1] at the step $20*5=100$, which shows that it has lost approx. $70$ states.

\begin{figure}[t]
\begin{minipage}{.45\textwidth}
\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/unr-events-new.png}
\end{minipage}
\hfill
%
\begin{minipage}{.45\textwidth}
\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/unr-events-old.png}
\end{minipage}
%
\caption{Dependency of number of events in the unrolled program on the unrolling bound}
\label{dep:events}
\end{figure}

@@

@@

%TO-DO: fix interface

%\textbf{TODOOOOOOOOOOOOOOOOOOOOO!!!! Re-compute tests with new counting of encoding}
The sample output of \porthos[2] in the state reachability analysis mode is presented in Appendix~\ref{apx:output}.
The output includes detailed time measurements for each stage of the analysis, and optionally statistical information about the unrolled program (number of events, transitions, etc.)

\begin{figure}[t]
\begin{minipage}{.45\textwidth}
\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/unr-time-new.png}
\end{minipage}
\hfill
%
\begin{minipage}{.45\textwidth}
\includegraphics[width=\textwidth,keepaspectratio]{img/my/performance/unr-time-old.png}
\end{minipage}
%
\caption{Dependency of program domain encoding time (in seconds) on the unrolling bound}
\label{dep:time}
\end{figure}

@@

Figure~\ref{dep:time} illustrates the dependency of overall time spent for the program domain encoding on the unrolling bound.
The performance of the encoding with the new unrolling algorithm used by \porthos[2] is approx. $1.6$ times worse than performance with the old unrolling scheme.

@@

@@

%<TODO: profile both programs, say sth about memory usage!>

@@

For evaluating the tool working in the portability analysis mode, we used the same file \texttt{Dekker.c} tested for the state-portability from TSO to SC: %TODO: Say sth about state-portability
%!!! \textit{\textbf{TODO: change the main class name to PorthosC }} !!!
As the portability analysis requires compiling the program under two memory models, the overall program domain encoding time 65.693 sec is almost double as the same time in reachability analysis mode (\porthos[1] shows program domain encoding time 41.027 sec with the bound $k_1=3$).
